---
execute:
  warning: false
  error: false
  cache: true
---

# `stormwindmodel` for Harvey

## Data Procurement

*   Hurricane track data was downloaded from the `hurricaneexposuredata` package using the `data("hurr_tracks")` command. This data originates from the [HURDAT2](https://www.nhc.noaa.gov/data/).
*   ZCTA to County Crosswalk was downloaded from <https://www2.census.gov/geo/docs/maps-data/data/rel/zcta_county_rel_10.txt>
*   ZCTA shapefiles were downloaded for the entire US using the `tigris` package

A vignette for the `hurricaneexposuredata` package can be found [here](https://cran.r-project.org/web/packages/hurricaneexposure/vignettes/hurricaneexposure.html)

## Data Processing

The `stormwindmodel` estimates wind exposure for points. In this case we used wind exposure at ZCTA geometric centroids to estimate wind exposure for the entire ZCTA.

1.  Hurricane data was downloaded using `hurricaneexposuredata` package using the `data("hurr_tracks")` command and filtered using the Harvey Automated Tropical Cyclone Forecasting code (AL092017)
2.  Separately, 2017 ZCTAs downloaded from the [US Census Cartographic Boundary Files](https://www.census.gov/geographies/mapping-files/time-series/geo/carto-boundary-file.2017.html#list-tab-1556094155)
3.  Then, the ZCTAs were joined to the [US Census ZCTA to County Crosswalk](https://www2.census.gov/geo/docs/maps-data/data/rel/zcta_county_rel_10.txt) in order to identify Texas and Louisiana ZCTAs
4.  The ZCTAs were filtered to Texas and Louisiana
5.  Geometrics centroids were calculated for each ZCTA
6.  The `get_grid_winds` command was used to calculate wind exposure for each centroid. It works by
    a.  Imputing the hurricane track to every 15 minutes
    b.  Adding the wind radii from the [HURDAT2](https://www.nhc.noaa.gov/data/)
    c.  Modeling wind exposure for each point

More information about the `stormwindmodel` package and variables calculated are included in this [vignette](https://cran.r-hub.io/web/packages/stormwindmodel/vignettes/Details.html)

## Output

-   `TXLA_ZCTA_AL092017_stormwindmodel.parquet`
    -   **GEOID** - ZCTA
    -   **date_time_max_wind** - date & time where maximum wind was sustained
    -   **vmax_sust** - Maximum 10-m 1-min sustained wind for the tropical cyclone (m/s)
    -   **vmax_gust** - Max 10-m 1-min gust wind experienced at grid point (m/s)
    -   **sust_dur** - Duration of time a certain sustained wind was experienced at grid point (minutes)
    -   **gust_dur** - Duration of time a certain gust wind was experienced at grid point (minutes)

## Code
Storm track data is filtered for Hurricane Harvey
```{r}
library(hurricaneexposuredata)
library(stormwindmodel)
library(tidyverse)

data("hurr_tracks")

AL092017_track <- hurr_tracks %>%
  filter(usa_atcf_id == "AL092017") %>%
  select(-c(storm_id, usa_atcf_id))
```

ZCTA geo-centoids were calculated
```{r}
#| output: false

library(tigris)
library(sf)

ZCTAs <- read_csv("https://www2.census.gov/geo/docs/maps-data/data/rel/zcta_county_rel_10.txt") %>%
  filter(STATE %in% c(48, 22)) %>%
  distinct(ZCTA5) %>%
  rename(GEOID = ZCTA5)

ZCTAs_sf <- ZCTAs %>%
  left_join(zctas(cb = FALSE, year = 2017), by = join_by(GEOID == GEOID10)) %>%
  st_as_sf() %>%
  select(-c(ZCTA5CE10, CLASSFP10, MTFCC10, FUNCSTAT10, ALAND10, AWATER10, INTPTLAT10, INTPTLON10))

ZCTAs_center <- ZCTAs_sf %>%
  st_transform("ESRI:102003") %>%
  st_centroid() %>%
  st_transform("EPSG:4269")

ZCTAs_centers_df <- ZCTAs_center %>%
  mutate(gridid = GEOID,
    glat = map_dbl(geometry, function(geo) st_coordinates(geo)[[2]]),
    glon = map_dbl(geometry, function(geo) st_coordinates(geo)[[1]]),
    glandsea = as.logical(1)) %>%
  select(gridid, glat, glon, glandsea) %>%
  st_drop_geometry() %>%
  filter(!is.na(glat))
```

Geo-centroids
```{r}
ggplot() + 
  geom_sf(data = ZCTAs_sf) + 
  geom_sf(data = ZCTAs_center, color = "red", size = 0.6) +
  coord_sf(crs = st_crs("ESRI:102003"))
```

Calculated wind exposure
```{r}
AL092017_winds <- get_grid_winds(hurr_track = AL092017_track, 
  grid_df = ZCTAs_centers_df) %>%
  rename(GEOID = gridid)
```

```{r}
AL092017_winds_sf <- AL092017_winds %>%
  left_join(zctas(cb = FALSE, year = 2017), by = join_by(GEOID == GEOID10)) %>%
  st_as_sf() %>%
  select(-c(ZCTA5CE10, CLASSFP10, MTFCC10, FUNCSTAT10, ALAND10, AWATER10, INTPTLAT10, INTPTLON10))
```

Plotted wind measures
```{r}
library(gridExtra)

output_maps <- c("vmax_sust", "vmax_gust", "sust_dur", "gust_dur") %>%
  map(~
    AL092017_winds_sf %>%
      ggplot() +
        geom_sf(aes(fill = !!sym(.x)), color = NA) +
        coord_sf(crs = st_crs("ESRI:102003"))
  )

do.call(grid.arrange, c(output_maps, ncol = 2))
```

```{r}
library(arrow)

AL092017_winds %>%
  write_parquet("output/TXLA_ZCTA_stormwindmodel.parquet")
```